---
title: "Untitled"
output: html_document
date: "2023-01-01"
---
## Variables
EFG_O: Effective Field Goal Percentage Shot
EFG_D: Effective Field Goal Percentage Allowed
TOR: Turnover Percentage Allowed (Turnover Rate)
TORD: Turnover Percentage Committed (Steal Rate)
ORB: Offensive Rebound Rate
DRB: Offensive Rebound Rate Allowed
FTR : Free Throw Rate (How often the given team shoots Free Throws)
FTRD: Free Throw Rate Allowed
2P_O: Two-Point Shooting Percentage
2P_D: Two-Point Shooting Percentage Allowed
3P_O: Three-Point Shooting Percentage
3P_D: Three-Point Shooting Percentage Allowed
ADJ_T: Adjusted Tempo (An estimate of the tempo (possessions per 40 minutes) a team would have against the team that wants to play at an average Division I tempo)
WAB: Wins Above Bubble (The bubble refers to the cut off between making the NCAA March Madness Tournament and not making it)

https://www.modb.pro/db/451190

```{r}
library(tidyverse)
library(finalfit)
library(ggsci)
library(ggpubr)
library(factoextra)
library(tidymodels)
library(janitor)
library(skimr)
library(modelr)
library(GGally)
library(kableExtra) # make nice looking results when we knit
library(vip)        # variable importance
library(fastshap)   # shapley values for variable importance 
library(MASS)
library(tree)
library(factoextra)
```

```{r}
ds <- read.csv('cbb.csv')
```


首先我想研究关于UNC这些年表现情况的情况，以及是什么因素影响了球队的表现。首先看下面这张图，所有红色的点就是UNC的数据。横轴是防守效率，即每百回合对手的得分，这个数据越小说明球队防守越好；而纵轴是进攻效率，即每百回合自己球队的得分，数据越大说明进攻越好。从这里我们可以看出来UNC的进攻效率和防守效率在所有球队中都是几乎最好的，因此我想研究球队具体是在哪方面表现得出色，以及在什么方面可以提升。

```{r}
pdf("Performance.jpg")
ggplot(ds,aes(ADJDE,ADJOE)) +
  geom_point() +
  geom_point(data = subset(ds, ds$TEAM == "North Carolina"),
             color = "red",
             size = 3) +
  theme_classic()
dev.off()
```

```{r}
summary(ds[which(ds$TEAM=='North Carolina'),c('ADJOE', 'ADJDE', 'WAB', 'EFG_O', 'EFG_D', 'TOR', 'TORD', 'DRB','ORB', 'FTR')])
```
```{r}
summary(ds[which(ds$POSTSEASON=="Champions"),c('ADJOE', 'ADJDE', 'WAB', 'EFG_O', 'EFG_D', 'TOR', 'TORD', 'DRB','ORB', 'FTR')])
```

```{r}
ds1 <- filter(ds, ds$TEAM=="North Carolina")
ds1
```

```{r}
summary(ds[which(ds$YEAR==2014),c('ADJOE', 'ADJDE', 'WAB', 'EFG_O', 'EFG_D', 'TOR', 'TORD', 'DRB','ORB', 'FTR')])
```


```{r}
m_adjoe <- lm(ADJOE~WAB+EFG_O+EFG_D+TOR+TORD+DRB+ORB+FTR, data=ds)
summary(m_adjoe)
```

```{r}
pdf("Relative.jpg")
df_adjoe <- data.frame(coefficients=c('WAB', 'EFG_O', 'EFG_D', 'TOR', 'TORD', 'DRB','ORB', 'FTR'), value=c(0.643662, 0.852157, 0.469380, -0.857523, -0.498525, 0.198763, 0.389387, 0.041582))

df_adjoe$coefficients <- factor(df_adjoe$coefficients, levels=df_adjoe$coefficients)

a<- ggplot(df_adjoe, aes(x=coefficients, y=value))+
  geom_bar(stat='identity', width=.5, fill="tomato3")+
  theme(axis.text.x = element_text(angle=65, vjust=0.6))+
  geom_text(aes(label=round(value,3)), 
            position = position_dodge2(width = 0.9, preserve = 'single'),
            vjust = -0.2, hjust = 0.5)+
  labs(title="Relevant Factors of Adjusted Offensive Efficiency", 
          caption="Source: cbb")

df_adjde <- data.frame(coefficients=c('WAB', 'X2P_D', 'X3P_D', 'DRB', 'ORB','TORD', 'ADJ_T'), value=c(-0.383978,0.798217,0.518325,-0.423676,0.071006,-0.801261,0.094749))

df_adjde$coefficients <- factor(df_adjde$coefficients, levels=df_adjde$coefficients)

b<- ggplot(df_adjde, aes(x=coefficients, y=value))+
  geom_bar(stat='identity', width=.5, fill="tomato3")+
  theme(axis.text.x = element_text(angle=65, vjust=0.6))+
  geom_text(aes(label=round(value,3)), 
            position = position_dodge2(width = 0.9, preserve = 'single'),
            vjust = -0.2, hjust = 0.5)+
  labs(title="Relevant Factors of Adjusted Defensive Efficiency", 
          caption="Source: cbb")

grid.arrange(a, b, ncol=2)
dev.off()
```








